### 2021-03-02 slides/sources/introduction.tex
re-organise layout of the slides

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The progress of AI and Robotics (AIR)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Savage N. 2021 in Nature}
\begin{frame}{The progress of AI and Robotics in USA, China and EU}
      \begin{figure}
        \centering
        %\includegraphics[width=1.0\textwidth]{./figures/air4children/versions/drawing-v02.png}
        %\caption{}
      \end{figure}
\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Open source software and hardware in AIR}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Savage N. 2021 in Nature}
\begin{frame}{Open source software and hardware in AIR}
      \begin{figure}
        \centering
        %\includegraphics[width=1.0\textwidth]{./figures/air4children/versions/drawing-v02.png}
        %\caption{}
      \end{figure}
\end{frame}
}