### 2021-03-02 slides/sources/conclusions-future-work.tex
#5 adds notes for images to be added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Lastname N. YEAR in journal of...}
\begin{frame}{Conclusions, future work and acknowledgements}

  \begin{columns}
    \begin{column}{.3\linewidth}
      Towards a more realistic
  \begin{itemize}
    \item \textbf{Fun}
    \item \textbf{Affortable}
    \item \textbf{Open Accessible}
    \item \textbf{D}
  \end{itemize}

    \end{column}

  \begin{column}{.6\linewidth}

      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/future-work/versions/drawing-v00.png}
      \end{figure}

    \end{column}
  \end{columns}

\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Lastname N. YEAR in journal of...}
\begin{frame}{Acknowledgements}

      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/future-work/versions/drawing-v00.png}
      \end{figure}

\end{frame}
}