### 2021-03-07 slides/sources/air4children.tex
polishes teaching section [skip-all]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{AIR4Children}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Prototyping and Piloting Open Source Robots}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Xochicale M. 2014 \it{Libre Robotics (proposal)} }
\begin{frame}{Prototyping Open Source Robots}
      \begin{figure}
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/air4children/versions/drawing-v00.png}
        %\caption{}
      \end{figure}
\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Xochicale M. 2015 in Mecate; Parra C. et al. 2016, \it{Otto DIY}}
\begin{frame}{Piloting robot prototypes}
      \begin{figure}
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/air4children/versions/drawing-v01.png}
        %\caption{}
      \end{figure}
\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Teaching Materials based on Montessori Education and Spiral Learning}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Elkin M., Sullivan A, Bers 2014 in Journal of Informantion and Technology}
\begin{frame}{AI and Robotics curriculum in Montessori Classroom}

  \it{The hand is the instrument of the mind.} Dr. Maria Montessori (1970-1952).
    \begin{figure}
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/montessori/versions/drawing-v00.png}
        %\caption{}
      \end{figure}
\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Mohammad T. et al., 2017; Harden R. M. and Stamper N in Journal of Medical Teacher}
\begin{frame}{Adopting Spiral Learning Methods in AIR4Children}

  \begin{figure}
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/teaching-materials/versions/drawing-v02.png}
        %\caption{}
      \end{figure}
\end{frame}
}