### 2021-03-07 slides/sources/conclusions-future-work.tex
polishes conclusions and future work [skip-all]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future work}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
%\paper{Lastname N. YEAR in journal of...}
\begin{frame}{Conclusions and future work}

  \begin{columns}
  \begin{column}{.75\linewidth}

  \textbf{Conclusions}

  We proposed AIR4children to create
  \begin{itemize}
    \item a more inclusive, affortable and fair participation of children in AI and Robotics
    \item child-centred AI and Robotics based on Spiral Learning and Montessori Education 
    \item Open source robots to be affortable and fun 
  \end{itemize}

  \textbf{Future work}
  \begin{itemize}
    \item Piloting air4children in Xicohtzinco, M\'exico
    \item Evaluation of the impact of AIR to children, instructors and parents. %(and perhaps their community).
  \end{itemize}

    \end{column}

  \begin{column}{.3\linewidth}

      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/future-work/versions/drawing-v00.png}
      \end{figure}

    \end{column}
  \end{columns}

\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
\paper{Lastname N. YEAR in journal of...}
\begin{frame}{Acknowledgements}

      \begin{figure}
        \centering
        \includegraphics[width=0.9\textwidth]{./figures/team/versions/drawing-v00.png}
      \end{figure}

\end{frame}
}