### 2021-03-08 slides/sources/introduction.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Education, Investment and Publications in AI and Robotics (AIR)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
%\paper{Savage N. 2021 in Nature, DOI: doi.org/10.1038/d41586-020-03409-8}
\paper{Zhang D. et al 2021 in AI 2021 Index Report, Fig 4.3.3 https://aiindex.stanford.edu/report/}

\begin{frame}{The progress of AI and Robotics in USA, China and EU}

\begin{figure}
 \centering
 \includegraphics[width=1.0\textwidth]{./figures/progress-of-air/versions/drawing-v00.png}
 %\caption{}
\end{figure}

\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
%\paper{Savage N. 2021 in Nature, DOI: doi.org/10.1038/d41586-020-03409-8}
\paper{Zhang D. et al 2021 in AI 2021 Index Report, Fig 1.1.19 https://aiindex.stanford.edu/report/}

\begin{frame}{The progress of AI and Robotics in USA, China and EU}

\begin{figure}
 \centering
 \includegraphics[width=1.0\textwidth]{./figures/progress-of-air/versions/drawing-v02.png}
 %\caption{}
\end{figure}

\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
%\paper{Savage N. 2021 in Nature, DOI: doi.org/10.1038/d41586-020-03409-8}
\paper{Zhang D. et al 2021 in AI 2021 Index Report, Fig 1.1.19 https://aiindex.stanford.edu/report/}

\begin{frame}{The progress of AI and Robotics in USA, China and EU}

\begin{figure}
 \centering
 \includegraphics[width=1.0\textwidth]{./figures/progress-of-air/versions/drawing-v01.png}
 %\caption{}
\end{figure}

\end{frame}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Open source software and hardware in AIR}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{
%\paper{Savage N. 2021 in Nature}
\begin{frame}{Open source software and hardware in AIR}

      \begin{figure}
        \centering
        \includegraphics[width=1.0\textwidth]{./figures/timeline-osh/versions/drawing-v00.png}
        %\caption{}
      \end{figure}
\end{frame}
}